---
title: "visualization_2"
author: "Yi"
date: "`r Sys.Date()`"
output: github_document
---

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

```{r}
library(tidyverse)
```

# load and clean data
```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```


# `group_by()`

```{r}
weather_df %>%
  group_by(name, month)
```

## `summarize()`

```{r}
weather_df %>%
  group_by(name) %>% 
  summarize()

weather_df %>%
  group_by(name) %>% 
  summarize(
    n_obs = n()
  )

weather_df %>%
  group_by(name, month) %>% 
  summarize(
    n_obs = n()
  )
```

additional way

```{r}
weather_df %>%
  count(month, name = "n_obs")
```

do not use `table()`, the result is not df

```{r}
weather_df %>%
  pull(month) %>% 
  table
```

multiple summaries

`distinc()`, default find distinct for whole data set.

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

(2x2 tables)

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ "" # default setting
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n())


weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ "" # default setting
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>%
  summarize(
    n_obs = n()
    ) %>% 
  pivot_wider(
    names_from = cold,
    values_from = n_obs
  )


weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ "" # default setting
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```

## General summaries

mean(), median(), var(), sd(), mad(), IQR(), min(), and max(). 

by default when calculate statistics if there is missing obs, then statistics would be missing.

use `na.rm = TRUE` to calculate mean ignore missing.

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))

weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax),
    median_tmax = median(tmax))

weather_df %>%
  group_by(name, month) %>%
  summarize(across(tmin:prcp, mean))
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax, na.rm=TRUE)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```

produce markdown table

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)
```

